A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species

### 2017-01-12 doc/paper1-submission/response.tex

Some elaboration on the sensitivity of their method to tree weights seems warranted. Moreover, the weights are appear to be applied for trees rather than nodes. I recognize that weighting nodes is likely to be impractical, but it could create problems if different source trees have different likelihoods of correctly resolving nodes at various depths in the tree, as seems likely to be the case (e.g., Chen et al. 2015). Perhaps the authors could elaborate on this issue.

Second, I am not sure that weights for trees can be obtained in a truly principled manner. I recognize that one could use any weighting scheme with the R\&H method but it would seem valuable to have an empirical example where a specific weighting scheme is used. Obviously, the weights used in opentree would be the most appropriate to examine. To evaluate those weights I found myself looking to Hinchliff et al. (2015) paper and the current tree on the opentree website. Overall, it would be good for readers to have as much information as possible about the weighting used by Hinchliff et al. (2015); I found the description of the weighting used by Hinchliff et al. (2015) difficult to understand. It would be useful for R\&H to provide as much additional information about Hinchliff et al. (2015) weights as possible, probably in the discussion.

With these concerns out of the way I will add that I am uncertain whether a supertree method that achieves “perfect” (or nearly perfect) weighting can be devised. The only method I can think of that might allow principled estimation of tree weights would be a version of the maximum likelihood supertree method described by Steel and Rodrigo (2008). However, I doubt that such an ML supertree method would meet all of the desiderata for a good supertree method that R\&H articulate (especially their first goal). It would be valuable for the authors to expand on the potential advantages (which appear to be largely computational) and pitfalls of their extreme weighting method.

\begin{reply}
2a. Concern: incorrect input trees could lead to a supertree that is not correct.  Response: our goal is to summarize the input trees, not to infer the true tree.  This would work just fine with incorrect input trees.

2b. Concern: hypotheses that occur in more input trees might not be correct, so a supertree method should not favor hypotheses just based on frequency.  Response: our method does not support hypotheses because of their frequency.  We use ranks instead.

2c. Concern: bad input trees could compromise the correctness of inferred phylogeny.  Downweighting bad input trees might assuage this.  Response: we are not performing a phylogeny inference, but attemptint to summarize input trees, so the correctness of the phylogeny is not an issue.  While it is possible to describe weights so that we can phrase our rank-based procedure as an optimization problem, at no point do we ever construct weights.

@@

2d. Concern: the weights have a large impact on the method.  Please describe how sensitive the method is to the weights.  Response: we indeed hope that the ranking has a substantial effect on the resulting tree.  If this were not the case, then ranks could not achieve their stated goal of allowing users to use ranks to fix problems in the tree.

Despite this, we ran the pipeline with the ranks reversed.  This drastic alteration of the ranks lead to a supertree that conflicted with 1264 branches of the original supertree.  Compared to this, the opentree 4.0 summary tree that used preceding methodology conflicts with 1387 branches of the 4.0’ tree using current methods.  However, most of the 36533 internal edges of the grafted tree are not altered.

2e. Concern: the weights are applied to trees rather than nodes.  This seems to be the only valid concern.

2f. Concern: the weights cannot be derived in a principled manner.  Results: we are not trying to estimate the weights from the input trees.  We would only need to estimate weights if we were trying to infer the true phylogeny from the input trees, a la the supertree versus supermatrix debate.  But we are not, therefore there is no need to derive weights.

2.  It seems that the reviewer thinks that we are trying to infer the true tree from a dataset consisting of the input trees, like Bininda-Edmonds.  Such an inference would need a principled method of resolving conflicts between input trees using only the information in the set of input trees.  The resulting supertree could then be used to critique the input trees, so that input trees that don't match the supertree would be considered to be incorrect (up to estimation error).

However, our approach is fundamentally different.  We do not seek to infer a tree that is more accurate than any individual input tree.  Instead we seek to summarize and describe the set of input trees.  While the supertree must conflict with some input trees, we do not claim that this constitutes an estimate of which input trees are correct or incorrect.  Instead we seek to represent all input trees (including those that conflict with the phylogeny estimate) by linking from summary tree edges to conflicting input tree edges via our annotations file.

Our approach does involve the resolution of conflict.  However, our conflict resolution is not phrased as an estimate of the true tree.  Instead, we supply rankings to allow curators to select which trees should win conflicts.

### 2017-01-17 doc/paper1-submission/response.tex

Also, a minor point with big implications: To prove a problem is NP-hard, a reduction *to* (not *from*) it is necessary. So, please replace "to Max-Clique" by "from Max-Clique" in Section 1.1.2.

\begin{reply}
Thanks - fixed.
\end{reply}

3. Performance: The authors chose one data set. Could they do an analysis of some additional data sets? For example, they can try the "perfect" case in which they take one of the supertrees they got, get out of it many subtrees (so, they know that these subtrees could be put back together into a supertree, and they know what that supertree is), and then run the methods?

\begin{reply}
Thanks for this suggestion.  We performed this experiment and added the following section describing the experiment and the results: \textcolor{red}{\textbf{Mark please check}}
\begin{quote}
To test the reliability of our methods, we divided the OTLv4' summary
tree into a collection of input trees. The input trees were obtained
by splitting the grafted tree at each taxonomy node, so that each
input tree has a taxonomy node at the root, and taxonomy nodes as
tips, but no taxonomy nodes internally.  Each leaf node that was an
internal taxon was then replaced with a leaf taxon. This approach
ensures that any taxa that were contested in the original analyses
remain contested in this new analysis.  We then constructed a summary
tree from these input trees using the propinquity pipeline.  The
resulting summary tree was identical to the OTLV4' summary tree.  
\end{quote}

Note that we are not trying to perform phylogeny estimation here like Bininda-Edmonds, so we are not treating the input trees as stand-ins for the data matrices that they may have been constructed from.  Because of this, we aren't treating the input tree collection as a data set.  To clarify this to the readers, we add the following text to the Background section:
\begin{quote}
Add non-goals section here (to be written).
\end{quote}

### 2017-01-17 doc/paper1-submission/paper5.tex
Update new section 3.4.

\subsection{Checking reliability of the method}
    To test the reliability of our methods, we divided the OpenTree 4' summary tree into a collection of input trees. The input trees were obtained by splitting the grafted tree at each taxonomy node, so that each input tree has a taxonomy node at the root, and taxonomy nodes as tips, but no taxonomy nodes internally. \emph{is this right?}  This approach ensures that any taxa that were contested in the original analyses remain contested in this new analysis.  We then constructed a summary tree from these input trees using the propinquity pipeline.  The resulting summary tree was identical to the OpenTree 4' summary tree.

\section{Conclusions}

>>>>

\subsection{Checking reliability of the method}
To test the reliability of our methods, we divided the OTLv4' summary
tree into a collection of input trees. The input trees were obtained
by splitting the grafted tree at each taxonomy node, so that each
input tree has a taxonomy node at the root, and taxonomy nodes as
tips, but no taxonomy nodes internally.  Each leaf node that was an
internal taxon was then replaced with a leaf taxon. This approach
ensures that any taxa that were contested in the original analyses
remain contested in this new analysis.  We then constructed a summary
tree from these input trees using the propinquity pipeline.  The
resulting summary tree was identical to the OTLV4' summary tree.  
\section{Conclusions}
